Quality Definitions of Qualities

### 2019-09-26 StateOfPractice/QDefOfQualities/QDefOfQualities.tex
Mostly empty document for adding our definitions of qualities

\title{Definining Quality} 
\author{Spencer Smith, Jacques Carette, Olu Owojaiye, Peter Michalski and Ao Dong}
\date{\today}

@@

Purpose and scope of the document.

\section{Qualities} \label{SecQualities}

To assess the impact of MDE on SCS quality, we need a clear definition of what
we mean by quality.  The concept of
quality is decomposed into a set of separate qualities.  This set of qualities
can be applied to the software artifacts (documentation, test cases, etc) and to
the software development process itself.  Several of the qualities from the list
in Section~\ref{SecSoftwareQuality} cannot be measured directly, such as
maintainability.  Therefore, Section~\ref{SecDesirableQs} introduces measurable
documentation qualities that are believed to contribute to the overall software
qualities.  In some cases it may be necessary to use the qualities in
Section~\ref{SecDesirableQs} to indirectly measure qualities listed in
Section~\ref{SecSoftwareQuality}.

@@

\subsection*{Installability}

@@

\item [\textbf{Correctness}] Software is correct if it matches its specification.

\item [\textbf{Verifiability}] involves ``solving the equations
  right''~\cite[p.~23]{Roache1998}; it benefits from rational documentation
  that systematically shows, with explicit traceability, how the governing
  equations are transformed into code.

\item [\textbf{Validatability}] means ``solving the right
  equations''~\cite[p.~23]{Roache1998}.  Validatability is improved by a
  rational process via clear documentation of the theory and assumptions, along
  with an explicit statement of the systematic steps required for experimental
  validation.

\item [\textbf{Reliability}] is a critical quality for scientific software,
  since the results of computations are meaningless, if they are not dependable.
  Reliability is closely tied to verifiability, since the key quality to verify
  is reliability, while the act of verification itself improves reliability.

\item [\textbf{Performance}] considerations can make certification challenging,
  since QA becomes more difficult for more complex code.  However, as
  Roache~\cite[p.~355]{Roache1998} points out, using simpler algorithms and
  reducing the number of options in general purpose code, is not always a
  practical option.

\item [\textbf{Usability}] can be a problem.  Different users, solving the same
  physical problem, using the same software, can come up with different answers,
  due to differences in parameter selection~\cite[p.~370]{Roache1998}.  To
  reduce misuse, a rational process must state expected user characteristics,
  modelling assumptions, definitions and the range of applicability of the code.

\item [\textbf{Maintainability}] is necessary in scientific software, since change,
  through iteration, experimentation and exploration, is inevitable.  Models of
  physical phenomena and numerical techniques necessarily evolve over
  time~\cite{CarverEtAl2007, SegalAndMorris2008}.  Proper documentation,
  designed with change in mind, can greatly assist with change management.%   QA
  % activities need to take the need for creativity into account, while not
  % smothering it~\cite[p.~352]{Roache1998}.

\item [\textbf{Reusability}] provides support for the quality of reliability,
  since reliability is improved by reusing trusted components~\cite{Dubois2005}.
  (Care must still be taken with reusing trusted components, since blind reuse
  in a new context can lead to errors, as dramatically shown in the Ariane 5
  disaster~\cite[p.~37--38]{OliveiraAndStewart2006}.)  The odds of reuse are
  improved when it is considered right from the start.

\item [\textbf{Understandability}] is necessary, since reviewers can only certify
  something they understand.  Scientific software developers have the
  view ``that the science a developer embeds in the code must be apparent to
  another scientist, even ten years later''~\cite{Kelly2013}.
  Understandability applies to the documentation and code, while usability
  refers to the executable software.  Documentation that follows a rational
  process is the easiest to follow.

\item [\textbf{Reproducibility}] is a required component of the scientific
  method~\cite{Davison2012}.  Although QA has, ``a bad name among
  creative scientists and engineers''~\cite[p.~352]{Roache1998}, the community
  need to recognize that participating in QA management also improves
  reproducibility.  Reproducibility, like QA, benefits from a consistent and
  repeatable computing environment, version control and separating code from
  configuration/parameters~\cite{Davison2012}.

\item [\textbf{Productivity}] \wss{This needs to be filled in.  Productivity is
    an important quality for this proposal, since part of what MDE promises is
    increased productivity.}

\item [\textbf{Sustainability}] \wss{This needs to be filled in.  A search needs
    to be done to find a good definition.  A starting point might be the recent
    paper that defines sustainability as a combination of other qualities.}

\end{description}

\subsection{Desirable Qualities of Documentation} \label{SecDesirableQs}

To achieve the qualities listed in Section~\ref{SecSoftwareQuality}, the
documentation should achieve the qualities listed in this section.  All but the
final quality listed (abstraction), are adapted from the IEEE recommended
practise for producing good software requirements~\cite{IEEE1998}.  Abstraction
means only revealing relevant details, which in a requirements document means
stating what is to be achieved, but remaining silent on how it is to be
achieved.  Abstraction is an important software development principle for
dealing with complexity~\cite[p.~40]{GhezziEtAl2003}.
\citet{SmithAndKoothoor2016} present further details on the qualities of
documentation for SCS.

\begin{description}

\item [\textbf{Completeness}] Documentation is said to be complete when all the
  requirements of the software are detailed. That is, each goal, functionality,
  attribute, design constraint, value, data, model, symbol, term (with its unit
  of measurement if applicable), abbreviation, acronym, assumption and
  performance requirement of the software is defined.  The software's response
  to all classes of inputs, both valid and invalid and for both desired and
  undesired events, also needs to be specified.

\item [\textbf{Consistency}] Documentation is said to be consistent when no subset
  of individual statements are in conflict with each other. That is, a
  specification of an item made at one place in the document should not
  contradict the specification of the same item at another location.

\item [\textbf{Modifiability}] The documentation should be developed in such a way
  that it is easily modifiable so that likely future changes do not destroy the
  structure of the document. Also it should be easy to reflect the change,
  wherever needed in the document to maintain consistency, traceability and
  completeness. For documentation to be modifiable, its format must be
  structured in a way that repetition is avoided and cross-referencing is
  employed.

\item [\textbf{Traceability}] Documentation should be traceable, as this
  facilitates maintenance and review. If a change is made to the design or code
  of the software, then all the documentation relating to those segments have to
  be modified.  This property is also important for recertification.

\item [\textbf{Unambiguity}] Documentation is said to be unambiguous only when
  every requirement's specification has a unique interpretation.  The
  documentation should be unambiguous to all audiences, including developers,
  users and reviewers.

\item [\textbf{Correctness}] There is no direct tool or method for measuring
  correctness. One way of building confidence in correctness is by reviewing to
  ensure that each requirement stated is one that the stakeholders and experts
  desire.  By maintaining traceability, consistency and unambiguity, we can
  reduce the occurrence of errors and make the goal of reviewing for correctness
  easier.

\item [\textbf{Verifiability}] Every requirement in the documentation must be the
  one fulfilled by the implemented software. Therefore all the requirements
  should be clear, unambiguous and testable, so that a person or a machine can
  verify whether the software product meets the requirements.

\item [\textbf{Abstract}] Documented requirements are said to be abstract if they
  state what the software must do and the properties it must possess, but do not
  speak about how these are to be achieved. For example, a requirement can
  specify that an Ordinary Differential Equation (ODE) must be solved, but it
  should not mention that Euler's method should be used to solve the ODE. How to
  accomplish the requirement is a design decision, which is documented during
  the design phase.

\end{description}

@@

\bibliographystyle {plainnat}
\bibliography {ResearchProposal}

>>>>

\title{Quality Definitions of Qualities} 
\author{Spencer Smith, Jacques Carette, Olu Owojaiye,\\Peter Michalski and Ao Dong}
\date{\today}

@@

Purpose and scope of the document.  \wss{Needs to be filled in.  Should
  reference the overal research proposal, and the ``state of the practice''
  exercise in particular.}

The presentation is divided into two main sections: i) qualities that apply to software products,
software artifacts and software development processes, and ii) qualities that
are considered important for good specifications.  The specification could be a
specification of requirements, design or a test plan.

\section{Qualities of Software Products, Artifacts and
  Processes} \label{SecQualities}

To assess the current state of software development, and to understand how
future changes impact software development, we need a clear definition of what
we mean by quality.  The concept of quality is decomposed into a set of separate
qualities.  This set of qualities can be applied to the software product, the
software artifacts (documentation, test cases, etc) and to the software
development process itself.

@@

\subsection{Installability}

@@

\subsection{Correctness}

Software is correct if it matches its specification.

\subsection{Verifiability}

Verifiability involves ``solving the equations right''~\cite[p.~23]{Roache1998};
it benefits from rational documentation that systematically shows, with explicit
traceability, how the governing equations are transformed into code.

\subsection{Validatability}

Validatability means ``solving the right equations''~\cite[p.~23]{Roache1998}.
Validatability is improved by a rational process via clear documentation of the
theory and assumptions, along with an explicit statement of the systematic steps
required for experimental validation.

\subsection{Reliability}

Reliability is a critical quality for scientific software, since the results of
computations are meaningless, if they are not dependable.  Reliability is
closely tied to verifiability, since the key quality to verify is reliability,
while the act of verification itself improves reliability.

\subsection{Robustness}

\subsection{Performance}

Performance considerations can make certification challenging, since QA becomes
more difficult for more complex code.  However, as
Roache~\cite[p.~355]{Roache1998} points out, using simpler algorithms and
reducing the number of options in general purpose code, is not always a
practical option.

\subsection{Usability}

Usability can be a problem.  Different users, solving the same physical problem,
using the same software, can come up with different answers, due to differences
in parameter selection~\cite[p.~370]{Roache1998}.  To reduce misuse, a rational
process must state expected user characteristics, modelling assumptions,
definitions and the range of applicability of the code.

\subsection{Maintainability}

Maintainability is necessary in scientific software, since change, through
iteration, experimentation and exploration, is inevitable.  Models of physical
phenomena and numerical techniques necessarily evolve over
time~\cite{CarverEtAl2007, SegalAndMorris2008}.  Proper documentation, designed
with change in mind, can greatly assist with change management.  QA activities
need to take the need for creativity into account, while not smothering
it~\cite[p.~352]{Roache1998}.

\subsection{Reusability}

Reusability provides support for the quality of reliability, since reliability
is improved by reusing trusted components~\cite{Dubois2005}.  (Care must still
be taken with reusing trusted components, since blind reuse in a new context can
lead to errors, as dramatically shown in the Ariane 5
disaster~\cite[p.~37--38]{OliveiraAndStewart2006}.)  The odds of reuse are
improved when it is considered right from the start.

\subsection{Portability}

\subsection{Understandability}

Understandability is necessary, since reviewers can only certify something they
understand.  Scientific software developers have the view ``that the science a
developer embeds in the code must be apparent to another scientist, even ten
years later''~\cite{Kelly2013}.  Understandability applies to the documentation
and code, while usability refers to the executable software.  Documentation that
follows a rational process is the easiest to follow.

\subsection{Interoperability}

\subsection{Visibility/Transparency}

\subsection{Reproducibility}

Reproducibility is a required component of the scientific
method~\cite{Davison2012}.  Although QA has, ``a bad name among creative
scientists and engineers''~\cite[p.~352]{Roache1998}, the community need to
recognize that participating in QA management also improves reproducibility.
Reproducibility, like QA, benefits from a consistent and repeatable computing
environment, version control and separating code from
configuration/parameters~\cite{Davison2012}.

\subsection{Productivity}

\subsection{Sustainability}

\section{Desirable Qualities of Good Specifications} \label{SecDesirableQs}

To achieve the qualities listed in Section~\ref{SecQualities}, the documentation
should achieve the qualities listed in this section.  All but the final quality
listed (abstraction), are adapted from the IEEE recommended practise for
producing good software requirements~\cite{IEEE1998}.  Abstraction means only
revealing relevant details, which in a requirements document means stating what
is to be achieved, but remaining silent on how it is to be achieved.
Abstraction is an important software development principle for dealing with
complexity~\cite[p.~40]{GhezziEtAl2003}.  \citet{SmithAndKoothoor2016} present
further details on the qualities of documentation for SCS.

\subsection{Completeness}

Documentation is said to be complete when all the requirements of the software
are detailed. That is, each goal, functionality, attribute, design constraint,
value, data, model, symbol, term (with its unit of measurement if applicable),
abbreviation, acronym, assumption and performance requirement of the software is
defined.  The software's response to all classes of inputs, both valid and
invalid and for both desired and undesired events, also needs to be specified.

\subsection{Consistency}

Documentation is said to be consistent when no subset of individual statements
are in conflict with each other. That is, a specification of an item made at one
place in the document should not contradict the specification of the same item
at another location.

\subsection{Modifiability}

The documentation should be developed in such a way that it is easily modifiable
so that likely future changes do not destroy the structure of the document. Also
it should be easy to reflect the change, wherever needed in the document to
maintain consistency, traceability and completeness. For documentation to be
modifiable, its format must be structured in a way that repetition is avoided
and cross-referencing is employed.

\subsection{Traceability}

Documentation should be traceable, as this facilitates maintenance and
review. If a change is made to the design or code of the software, then all the
documentation relating to those segments have to be modified.  This property is
also important for recertification.

\subsection{Unambiguity}

Documentation is said to be unambiguous only when every requirement's
specification has a unique interpretation.  The documentation should be
unambiguous to all audiences, including developers, users and reviewers.

\subsection{Correctness}

There is no direct tool or method for measuring correctness. One way of building
confidence in correctness is by reviewing to ensure that each requirement stated
is one that the stakeholders and experts desire.  By maintaining traceability,
consistency and unambiguity, we can reduce the occurrence of errors and make the
goal of reviewing for correctness easier.

\subsection{Verifiability}

Every requirement in the documentation must be the one fulfilled by the
implemented software. Therefore all the requirements should be clear,
unambiguous and testable, so that a person or a machine can verify whether the
software product meets the requirements.

\subsection{Abstract}

Documented requirements are said to be abstract if they state what the software
must do and the properties it must possess, but do not speak about how these are
to be achieved. For example, a requirement can specify that an Ordinary
Differential Equation (ODE) must be solved, but it should not mention that
Euler's method should be used to solve the ODE. How to accomplish the
requirement is a design decision, which is documented during the design phase.

@@

\bibliographystyle {plainnat}
\bibliography {../../CommonFiles/ResearchProposal}
